---
title: "Practical Machine Learning Course Project"
author: "Tan Woei Ming"
date: "27 December 2015"
output: html_document
---


# __A. Goal:__
To predict the manner in which they did the exercise, which is the "classe" variable in the training set.

# __B. R code and steps to build/train a model:__
##__B.1 Data exploration__
Include the R library for the machine learning
```{r }
library(caret)
library(useful)
```

Read in the training data, which consist of 19622 records, with 160 variables, including the "classe"s.
```{r }
datatrain = read.csv("pml-training.csv")
```

Data exploration to understand the each variable, or feature in the model. Remove the features with NA/NULL from the model building
```{r summary(datatrain)}
summary(datatrain)
```

##__B.2 Create the train and cross validation data set__
Create the train (75%) and cross validation (25%) data set, called datatrain_clean and datacv_clean respectively.
```{r }
inTrain<-createDataPartition(y=datatrain$classe,p=0.75,list=FALSE)
datatrain_clean<-datatrain[inTrain,8:160]
datacv_clean<-datatrain[-inTrain,8:160]
```

Load the randomForest library and set the seed to 9747
```{r }
library(useful)
library(randomForest)
set.seed(9747)
```

Create the parameters for building for x and y in randomForest using the train data. Based on the data exploration, select the feature for model building
```{r }
rfdata<-classe~roll_belt+pitch_belt+yaw_belt+total_accel_belt+gyros_belt_x+gyros_belt_y+gyros_belt_z+accel_belt_x+accel_belt_y+accel_belt_z+magnet_belt_x+magnet_belt_y+magnet_belt_z+roll_arm+pitch_arm+yaw_arm+total_accel_arm+gyros_arm_x+gyros_arm_y+gyros_arm_z+accel_arm_x+accel_arm_y+accel_arm_z+magnet_arm_x+magnet_arm_y+magnet_arm_z+roll_dumbbell+pitch_dumbbell+yaw_dumbbell+total_accel_dumbbell+gyros_dumbbell_x+gyros_dumbbell_y+gyros_dumbbell_z+accel_dumbbell_x+accel_dumbbell_y+accel_dumbbell_z+magnet_dumbbell_x+magnet_dumbbell_y+magnet_dumbbell_z+roll_forearm+pitch_forearm+yaw_forearm+total_accel_forearm+gyros_forearm_x+gyros_forearm_y+gyros_forearm_z+accel_forearm_x+accel_forearm_y+accel_forearm_z+magnet_forearm_x+magnet_forearm_y+magnet_forearm_z
xrf<-build.x(rfdata,data=datatrain_clean)
yrf<-build.y(rfdata,data=datatrain_clean)
```

The response and selected features:
```{r}
rfdata<-classe~roll_belt+pitch_belt+yaw_belt+total_accel_belt+gyros_belt_x+gyros_belt_y+gyros_belt_z+accel_belt_x+accel_belt_y+accel_belt_z+magnet_belt_x+magnet_belt_y+magnet_belt_z+roll_arm+pitch_arm+yaw_arm+total_accel_arm+gyros_arm_x+gyros_arm_y+gyros_arm_z+accel_arm_x+accel_arm_y+accel_arm_z+magnet_arm_x+magnet_arm_y+magnet_arm_z+roll_dumbbell+pitch_dumbbell+yaw_dumbbell+total_accel_dumbbell+gyros_dumbbell_x+gyros_dumbbell_y+gyros_dumbbell_z+accel_dumbbell_x+accel_dumbbell_y+accel_dumbbell_z+magnet_dumbbell_x+magnet_dumbbell_y+magnet_dumbbell_z+roll_forearm+pitch_forearm+yaw_forearm+total_accel_forearm+gyros_forearm_x+gyros_forearm_y+gyros_forearm_z+accel_forearm_x+accel_forearm_y+accel_forearm_z+magnet_forearm_x+magnet_forearm_y+magnet_forearm_z
```

Create the parameters for building for xtest and ytest in randomForest using the cross validation data.
```{r }
rfdatapred<-classe~roll_belt+pitch_belt+yaw_belt+total_accel_belt+gyros_belt_x+gyros_belt_y+gyros_belt_z+accel_belt_x+accel_belt_y+accel_belt_z+magnet_belt_x+magnet_belt_y+magnet_belt_z+roll_arm+pitch_arm+yaw_arm+total_accel_arm+gyros_arm_x+gyros_arm_y+gyros_arm_z+accel_arm_x+accel_arm_y+accel_arm_z+magnet_arm_x+magnet_arm_y+magnet_arm_z+roll_dumbbell+pitch_dumbbell+yaw_dumbbell+total_accel_dumbbell+gyros_dumbbell_x+gyros_dumbbell_y+gyros_dumbbell_z+accel_dumbbell_x+accel_dumbbell_y+accel_dumbbell_z+magnet_dumbbell_x+magnet_dumbbell_y+magnet_dumbbell_z+roll_forearm+pitch_forearm+yaw_forearm+total_accel_forearm+gyros_forearm_x+gyros_forearm_y+gyros_forearm_z+accel_forearm_x+accel_forearm_y+accel_forearm_z+magnet_forearm_x+magnet_forearm_y+magnet_forearm_z
xrfpred<-build.x(rfdatapred,data=datacv_clean)
yrfpred<-build.y(rfdatapred,data=datacv_clean)
```

##__B.3 Train the model__
Set x and y using the train set, and xtest and ytest using the cross validation set. Set all other parameters as default.
```{r}
rfrtrain<-randomForest(x=xrf,y=yrf,xtest = xrfpred,ytest=yrfpred)
rfrtrain
```
__The model could achieve >99% accuracy and sample error of 0.49%.__

# __C. Predict the test data__
Load the test data
```{r }
datatest = read.csv("pml-testing.csv")
```

Select the features for model fitting, the same features used by the training and cross validation. Also, add a dummy factor variable factor(c("A","B","C","D","E")) called "classe" for randomForest fitting.
```{r}
datatest_clean<-datatest[,8:159]
datatest_clean[,"classe"]<-factor(c("A","B","C","D","E"))
datatest_clean
```

Fit the test data into the model and predict the result. Extract the output:
```{r }
rfdatatest<-classe~roll_belt+pitch_belt+yaw_belt+total_accel_belt+gyros_belt_x+gyros_belt_y+gyros_belt_z+accel_belt_x+accel_belt_y+accel_belt_z+magnet_belt_x+magnet_belt_y+magnet_belt_z+roll_arm+pitch_arm+yaw_arm+total_accel_arm+gyros_arm_x+gyros_arm_y+gyros_arm_z+accel_arm_x+accel_arm_y+accel_arm_z+magnet_arm_x+magnet_arm_y+magnet_arm_z+roll_dumbbell+pitch_dumbbell+yaw_dumbbell+total_accel_dumbbell+gyros_dumbbell_x+gyros_dumbbell_y+gyros_dumbbell_z+accel_dumbbell_x+accel_dumbbell_y+accel_dumbbell_z+magnet_dumbbell_x+magnet_dumbbell_y+magnet_dumbbell_z+roll_forearm+pitch_forearm+yaw_forearm+total_accel_forearm+gyros_forearm_x+gyros_forearm_y+gyros_forearm_z+accel_forearm_x+accel_forearm_y+accel_forearm_z+magnet_forearm_x+magnet_forearm_y+magnet_forearm_z
xrftest<-build.x(rfdatatest,data=datatest_clean)
yrftest<-build.y(rfdatatest,data=datatest_clean)
rfresult<-randomForest(x=xrf,y=yrf,xtest = xrftest,ytest=yrftest)
rfresult
```

```{r}
rfpred<-rfresult["test"][[1]][1]
rfpred
```

# __D. Conclusion__
The randomForest model to train and predict "classe" is built and able to predict >99% accuracy. The model is also used to predict the 20 test case in the assignment.
Path to repository:<https://github.com/twming/practicalmachinelearning/>

# __--End of Report--__

# __Appendix__
Other models:
LDA give ~70% accurary
```{r }
#train the model using lda
set.seed(9747)
mylda=train(classe~roll_belt+pitch_belt+yaw_belt+total_accel_belt+gyros_belt_x+gyros_belt_y+gyros_belt_z+accel_belt_x+accel_belt_y+accel_belt_z+magnet_belt_x+magnet_belt_y+magnet_belt_z+roll_arm+pitch_arm+yaw_arm+total_accel_arm+gyros_arm_x+gyros_arm_y+gyros_arm_z+accel_arm_x+accel_arm_y+accel_arm_z+magnet_arm_x+magnet_arm_y+magnet_arm_z+roll_dumbbell+pitch_dumbbell+yaw_dumbbell+total_accel_dumbbell+gyros_dumbbell_x+gyros_dumbbell_y+gyros_dumbbell_z+accel_dumbbell_x+accel_dumbbell_y+accel_dumbbell_z+magnet_dumbbell_x+magnet_dumbbell_y+magnet_dumbbell_z+roll_forearm+pitch_forearm+yaw_forearm+total_accel_forearm+gyros_forearm_x+gyros_forearm_y+gyros_forearm_z+accel_forearm_x+accel_forearm_y+accel_forearm_z+magnet_forearm_x+magnet_forearm_y+magnet_forearm_z,data=datatrain_clean,method="lda")

#verify the model using the cross-validation dataset
mylda$finalModel
mypred<-predict(mylda,newdata=datacv_clean)
confusionMatrix(mypred,datacv_clean$classe)
table(mypred,datacv_clean$classe)
```

